---
title: "novel-predictions"
author: "George G Vega Yon"
date: "9/24/2019"
output:
  beamer_presentation:
    keep_tex: true
header-includes:
  - \usepackage{booktabs}
  - \usepackage{tabularx}
  - \usepackage{geometry}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(aphylo)
library(xtable)
```

## Estimates

```{r read-data}
files <- list.files(pattern = "parameter_estimates.+rds", full.names = FALSE)
estimates <- lapply(files, readRDS)
```

```{r coef, results='asis'}
theta <- do.call(cbind, lapply(estimates, coef))
vars  <- do.call(cbind, lapply(lapply(estimates, vcov), diag))

tab  <- matrix(sprintf("%.2f", theta), ncol = ncol(theta))
dimnames(tab) <- list(rownames(theta), paste0("(", 1:ncol(tab), ")"))

varnames <- c(
  psi0  = "$\\psi_0$",
  psi1  = "$\\psi_1$",
  mu_d0 = "$\\mu_{d0}$",
  mu_d1 = "$\\mu_{d1}$",
  mu_s0 = "$\\mu_{s0}$",
  mu_s1 = "$\\mu_{s1}$",
  Pi    = "$\\pi$"
  )

rownames(tab) <- varnames[rownames(tab)]

# Adding information

ps <- lapply(estimates, prediction_score)
aucs <- sapply(ps, function(i) {
  mean(sapply(i, function(j) aphylo::auc(j$predicted, j$expected)$auc))
})
pss  <- sapply(ps, function(i) {
  tmp <- 1 - t(sapply(i, function(j) with(j, c(obs, random)/worse)))
  colMeans(tmp, na.rm = TRUE)
})


tab <- rbind(
  tab,
  `Tree count` = sapply(estimates, Ntrees),
  Method       = ifelse(grepl("mcmc", files), "MCMC", "MLE"),
  Prior        = ifelse(grepl("no_prior", files), "Uniform", "Beta"),
  Inferred     = ifelse(grepl("fully", files), "Yes", "No"),
  AUC          = sprintf("%.2f", aucs),
  `P. Score (obs)` = sprintf("%.2f", pss[1,]),
  `P. Score (random)` = sprintf("%.2f", pss[2,])
  )

tmptab <- tempfile()
tab <- xtable(tab, caption = paste0(
  "Parameter estimates using different estimation methods, priors, and ",
  "types of annotations."
  ))
print(tab, file = tmptab, booktabs = TRUE, scalebox = .6, 
      comment = FALSE, sanitize.text.function= function(i) i)
tab <- readLines(tmptab)
tab[grepl("begin\\{tabular", tab)] <- sprintf(
  "\\begin{tabular}{l*{%d}{m{%.2f\\linewidth}}}",
  ncol(theta),
  1.1/ncol(theta)
)
tab[grepl("Method", tab)] <- paste("\\midrule", tab[grepl("Method", tab)])

cat(tab, sep="\n")
```

## MCMC convergence: Fully annotated (Unif prior)

```{r preparing-for-mcmc-diagnostics}
files <- gsub("(parameter_estimates_)", "", files)
files <- gsub("(mcmc|mle)", "\\U\\1", files, perl = TRUE)
files <- gsub("_", " ", files)
files <- gsub("[.]rds", "", files)
names(estimates) <- files

# Getting the 4 smallest trees
inferred <- order(Ntip(estimates$`MCMC fully annotated no prior`))[1:4]
observed <- order(Ntip(estimates$`MCMC partially annotated no prior`))[1:4]
```


```{r setting-coda}
library(coda)
traceplot2 <- function(x, ylim = c(0,1), what = NULL) {
  x <- x$hist
  
  if (!is.null(what))
    x <- what(x)
  
  k <- ifelse(is.mcmc.list(x), ncol(x[[1]]), ncol(x))
  op <- par(mfrow = c(4,2), mai = c(0.3366, 0.2706, 0.2706, 0))
  traceplot(x, ylim = ylim)
  par(op)
}

```


```{r trace-fully-unif}
traceplot2(estimates$`MCMC fully annotated no prior`)
```

## MCMC convergence: Fully annotated (Beta prior)

```{r trace-fully-beta}
traceplot2(estimates$`MCMC fully annotated prior`)
```

## MCMC convergence: Partially annotated (Unif prior)

```{r trace-partially-unif}
traceplot2(estimates$`MCMC partially annotated no prior`)
```

## MCMC convergence: Partially annotated (Beta prior)

```{r trace-partially-beta}
traceplot2(estimates$`MCMC partially annotated prior`)
```

---

Zooming in:

```{r}
traceplot2(estimates$`MCMC partially annotated prior`, ylim=NULL,
           what = function(i) window(i, 2000))
```


## How priors affect predictions: Fully annotated (Unif prior)

```{r annotations1.1}
plot(estimates$`MCMC fully annotated no prior`, which.tree = inferred[2])
```

## How priors affect predictions: Fully annotated (Beta prior)

```{r annotations1.2}
plot(estimates$`MCMC fully annotated prior`, which.tree = inferred[2])
```


## How priors affect predictions: Experimentally annotated (Unif prior)

```{r annotations2.1}
plot(estimates$`MCMC partially annotated no prior`, which.tree = observed[1])
```


## How priors affect predictions: Experimentally annotated (Beta prior)

```{r annotations2.2}
plot(estimates$`MCMC partially annotated prior`, which.tree = observed[1])
```

