---
title: "P2 Simulation Results"
output:
  beamer_presentation:
    slide_level: 2
author: George G Vega Yon
date: Sep 27, 2018
header-includes: \usepackage{booktabs}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Intro

*   Took the first 2,000 trees from panther

*   Ran 4 types of models:
    1.  Gold Standard (no missing) `dat ~ psi + eta + mu + Pi`
    2.  Missing `dat ~ psi + eta + mu + Pi`
    3.  Pub Bias (has missing and publication bias) `dat ~ mu + psi + Pi`
    4.  Full model (same as before) `dat ~ psi + eta + mu + Pi`
    
*   Each model ran the MCMCs with the following parameteres:
    
    ```r
    mcmc.nbatch  <- 1e5
    mcmc.burnin  <- 2e4
    mcmc.thin    <- 100
    mcmc.nchains <- 4
    ```
    
    Moreover, each chain now starts from a randomly chosen point.

# Bias

## Contents

\tableofcontents[currentsection]

## 01-gold-standard: Small

```{r, out.width=".6\\linewidth", fig.align='center'}
knitr::include_graphics("01-gold-standard/bias_plots_tree-size=small.pdf")
```

## 01-gold-standard: Mid-small

```{r, out.width=".6\\linewidth", fig.align='center'}
knitr::include_graphics("01-gold-standard/bias_plots_tree-size=mid-small.pdf")
```

## 01-gold-standard: Mid-large

```{r, out.width=".6\\linewidth", fig.align='center'}
knitr::include_graphics("01-gold-standard/bias_plots_tree-size=mid-large.pdf")
```

## 01-gold-standard: Large

```{r, out.width=".6\\linewidth", fig.align='center'}
knitr::include_graphics("01-gold-standard/bias_plots_tree-size=large.pdf")
```

## 04-full-model: Small

```{r, out.width=".6\\linewidth", fig.align='center'}
knitr::include_graphics("04-full-model/bias_plots_tree-size=small.pdf")
```

## 04-full-model: Mid-small

```{r, out.width=".6\\linewidth", fig.align='center'}
knitr::include_graphics("04-full-model/bias_plots_tree-size=mid-small.pdf")
```

## 04-full-model: Mid-large

```{r, out.width=".6\\linewidth", fig.align='center'}
knitr::include_graphics("04-full-model/bias_plots_tree-size=mid-large.pdf")
```

## 04-full-model: Large

```{r, out.width=".6\\linewidth", fig.align='center'}
knitr::include_graphics("04-full-model/bias_plots_tree-size=large.pdf")
```

# Convergence

## Contents

\tableofcontents[currentsection]

## 01-gold-standard

```{r, fig.align='center', out.width=".7\\linewidth"}
knitr::include_graphics("01-gold-standard/gelmans.pdf")
```


## 02-missing

```{r, fig.align='center', out.width=".7\\linewidth"}
knitr::include_graphics("02-missing/gelmans.pdf")
```


## 03-pub-bias

```{r, fig.align='center', out.width=".7\\linewidth"}
knitr::include_graphics("03-pub-bias/gelmans.pdf")
```


## 04-full-model

```{r, fig.align='center', out.width=".7\\linewidth"}
knitr::include_graphics("04-full-model/gelmans.pdf")
```


# Coverage

## Contents

\tableofcontents[currentsection]

## 01-gold-standard

\input{../tables/01_coverage95_by_method_missingness.tex}

## 02-missing

\input{../tables/02_coverage95_by_method_missingness.tex}

## 03-pub-bias

\input{../tables/03_coverage95_by_method_missingness.tex}


## 04-full-model

\input{../tables/04_coverage95_by_method_missingness.tex}


# Prediction

## Contents

\tableofcontents[currentsection]

## 01-gold-standard

```{r, fig.align='center', out.width=".7\\linewidth"}
knitr::include_graphics("01-gold-standard/mcmc_right_prior_prediction.pdf")
```


## 02-missing

```{r, fig.align='center', out.width=".7\\linewidth"}
knitr::include_graphics("02-missing/mcmc_right_prior_prediction.pdf")
```


## 03-pub-bias

```{r, fig.align='center', out.width=".7\\linewidth"}
knitr::include_graphics("03-pub-bias/mcmc_right_prior_prediction.pdf")
```


## 04-full-model

```{r, fig.align='center', out.width=".7\\linewidth"}
knitr::include_graphics("04-full-model/mcmc_right_prior_prediction.pdf")
```

